---
title: "Week 3 Exercises"
author: "HD Sheets"
date: "10/1/2024"
output:
  html_document:
    df_print: paged
student: James Mahar submitted 11/10/2024
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{R}
library('tidyverse')
```


Please complete all exercises below. You may use any library that we have covered in class UP TO THIS POINT. 

For each function,  show that it works, by using the provided data as a test
and by feeding in some test data that you create to test your function

Add comments to your function to explain what each line is doing

#Problem 1 
1.) Write a function that takes in a string with a person's name in the form

"Sheets, Dave"

and returns a string of the form

"Dave Sheets"

Note:

-assume no middle initial ever
-remove the comma
-be sure there is white space between the first and last name

You will probably want to use stringr

```{R}
name_in = "Sheets, Dave"

reorder_name <- function(name_in) 
#split name_in into first and last names by splitting the name_in string at the comma delimiter
{ name_parts <- str_split(name_in, ', ')[[1]] 
#first_name <- name_parts[2] #find first name
#last_name <- name_parts[1]  #find last nnme
#swapped_name <- paste(first_name, last_name) #reordered name as first space last
}
print(name_in) #print input
#print(swapped_name) #print 
#print(first_name) #print name parts
#print(last_name)
#print("From",name_in, "To",swapped_name)
```
#The [1] "Bob"   "Jane"  "Angel" output is from an old df. Why is it printing here? Plus, "swapped_name <-" does not run


#Problem 2 ???
2.) Write a function that takes in a string of values x, and returns a data frame
with three columns,  x, x^2 and the square root of x

```{R}
#Steps: name function, split the x string and convert to a vector, create a dataframe of 3 columns of x, x^2, x^0.5

x=c(1,3,5,7,9,11,13) #setup string of values to work with
powers_df<-function(x) #create function 
{vec <- unlist(strsplit(x, ",")) #split the input string of values minus comma delimiters
df <- data.frame(             #create a dataframe with calculated columns 
original = as.numeric(vec), 
square = as.numeric(vec)^2,
sqroot = as.numeric(vec)^.5
)
}
print(x) #print input string 
print(powers_df) #print the function
print(df) #print the dataframe of calculated column
```
#I know this output is wrong. I am confused. I reassigned df so, why am I seeing output from an old df?

#Problem 3 
3.) Write in a function that takes in a value x and returns
    
    y= 0.3x if x<0
    y=0.5x if x>=0
    
    This is a variant on a relu function as used in some neural networks.

```{r}

#prob3func <- function(x) #create function... 
#{
#y=ifelse(x<7,0.3*x,0.5*x). #using ifelse() command
#}
#print(prob3func(2)). #print function output changing the value of x to check the code for correctness

```

#Problem 4 
4.) Write a function that takes in a value x and returns the first power of two
greater than x  (use a While loop)
#Unfortunately I am stuck on this one!
```{r}

#prob4func <- function(x) {
#calculate 2^1,2^2,2^3...2^y compare to x and return first 2^y > x
#while (x<2^y)
#}

```

#Problem 5 
5) Two Sum - Write a function named two_sum()

Given a vector of integers nums and an integer target, return indices of the two numbers such that they add up to target.

You may assume that each input would have exactly one solution, and you may not use the same element twice.

You can return the answer in any order.

Example 1:

Input: nums = [2,7,11,15], target = 9
Output: [0,1]
Explanation: Because nums[0] + nums[1] == 9, we return [0, 1].
Example 2:

Input: nums = [3,2,4], target = 6
Output: [1,2]
Example 3:

Input: nums = [3,3], target = 6
Output: [0,1]

Constraints:

2 <= nums.length <= 104
–109 <= nums[i] <= 109
–109 <= target <= 109
Only one valid answer exists.
 

*Note: For the first problem I want you to use a brute force approach (loop inside a loop)* 

*The brute force approach is simple. Loop through each element x and find if there is another value that equals to target – x*

*Use the function seq_along to iterate*

```{r}
two_sum <- function(nums_vector,target){
    for(i in seq_along(nums_vector)) {
    for(j in seq_along(nums_vector)[-length(nums_vector)]) {
      
      sum <- nums_vector[i] + nums_vector[j + 1]
      if(sum == target) {
        
        first <- i
        second <- j + 1
        indexes <- c(first, second)
        return(indexes)
        
      }
      
    }
    
  }
  
}
# Test code
nums_vector <- c(5,7,12,34,6,10,8,9)
target <- 13
 
z=two_sum(nums_vector,target)
print(z)
#expected answers
#[1] 1 7
#[1] 2 5
#[1] 5 2

```

#Problem 6 
6.) Write one piece of code that will use a regex command to extract a phone number written in the form

      456-123-2329
      
The sentences to use are located below

use the str_extract function from stringr

use the same regex search pattern for each

-What does \\d match to?  or alternatively [[:digit:]] ?
Ans: I do not understand the question

-How do you specify a specific number of repeated characters ?       
Ans: "4{3,}" specifies 3 of the the number 4 in the str_detect command: str_detect(col1, "4{3,}") 

```{R}

a="Please call me at 456-123-2329, asap"

b="Hey, we have a code 234 on machine a-234-12, call me at 678-321-98766"

c="On 12-23-2022, Joe over at 122 Turnpike, dialled 912-835-4756, tell me by 9:02 pm Wed"


#Look for the following pattern in each variable, a,b,c: \\d{3}(-)\\d{3}(-)\\d{4} and print the number found

aphone_number <- str_extract(a,"\\d{3}(-)\\d{3}(-)\\d{4}") #extract and print the string pattern specified
print(aphone_number)

bphone_number <- str_extract(b,"\\d{3}(-)\\d{3}(-)\\d{4}") #extract and print the string pattern specified
print(bphone_number)

cphone_number <- str_extract(c,"\\d{3}(-)\\d{3}(-)\\d{4}") #extract andf print the string pattern specified
print(cphone_number)

```

#Problem 7 
7.)  For lines below,  extract the domains (ie the part of the address after @)

```{R}

d="jimmy.halibut@gmail.com"

e="His address is:  c.brown@hopeles.org, do write him"

f="h.potter@hogwarts.edu is bouncing back on me, I wonder why?"

string <- c("jimmy.halibut@gmail.com","His address is:  c.brown@hopeles.org, do write him","h.potter@hogwarts.edu is bouncing back on me, I wonder why?" )


str_split(d, pattern = "@") %>%  
map_chr(2)


dummy <- str_split(e, pattern = "@") %>% 
map_chr(2)

str_split(dummy, pattern = ", ") %>%
map_chr(1)

dumdum <- str_split(f, pattern = "@") %>% 
map_chr(2)

str_split(dumdum, pattern = " ") %>%
map_chr(1)

#library(urltools)
#suffix_extract(domain(string))

```

